---
title: "Writing Functions"
subtitle: "Module 1: Reproducible Research Example"
author: "Dr Alyce Russell <br> Postdoctoral Research Fellow <br> Centre for Precision Health, ECU <br> a.russell@ecu.edu.au <br> https://www.twitter.com/nerdrusty "
date: "Created: 3 March 2021 <br> Last Updated: `r format(Sys.Date(), format = '%d %B %Y')`"
output: 
  html_document:
    css: buttons.css
    fig_caption: yes
    highlight: tango
    keep_md: yes
    theme: lumen
    toc: yes
    toc_float:
      collapsed: false
    toc_depth: 2
    self_contained: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

```{r, echo=FALSE}
#Determine the output format of the document
outputFormat = knitr::opts_knit$get("rmarkdown.pandoc.to")

#Figure and Table Caption Numbering, for HTML do it manually
capTabNo = 1; capFigNo = 1;

#Function to add the Table Number
capTab = function(x){
  if(outputFormat == 'html'){
    x = paste0("**", "Table ", capTabNo, ". ", "**", x)
    capTabNo <<- capTabNo + 1
  }; x
}

#Function to add the Figure Number
capFig = function(x){
  if(outputFormat == 'html'){
    x = paste0("**", "Figure ",capFigNo,". ", "**", x)
    capFigNo <<- capFigNo + 1
  }; x
}

```

<div style="margin-bottom:50px;">

<style>
 #TOC {
  background: url("./images/Centre-for-Precision-Health.png");
  background-size: contain;
  padding-top: 100px !important;
  background-repeat: no-repeat;
}
</style>

</div>

<style>
  p.caption {
    font-size: 1.2em;
    }
  caption {
    font-size: 1.2em;
    } 
  body{
    font-size: 12pt;
    }
</style>


<script src="hideOutput.js"></script>   


## Overview

<br>

This is a walk-through on writing functions in R. We will use publicly available data to complete this task (downloaded from [here](http://faculty.washington.edu/kenrice/bigr/index.shtml) but also available with these module 1 notes. 

<br>

```{r}
library(tidyverse)
library(readxl)
library(kableExtra)
library(ggpubr)
```

## Data Preparation

<br>

The data are some SNPs and basic phenotypes for 1000 individuals. 

```{r}
snps <- read.csv("./data/example-snp.csv", header = TRUE)
pheno <- read.csv("./data/example-pheno.csv", header = TRUE)
```

<br>

Let's explore the data a little bit. 

<br>

```{r}
str(snps)

str(pheno)
```

<br>

## Prepare Data

<br>

Let's write a function to determine the minor allele, minor allele frequency in this population, and convert data to numeric coding (0,1,2) with 2 indicating the minor allele.  

<br>

```{r}
source("./rscripts/cleansnps.R")

snps_coded <- cleansnps(snps, id_var="id")
snpsbim <- cleansnps(snps, id_var="id", output="snpsbim")
```


